---
title: "Homework 2, P8105"
author: "Lingke (Lincole) Jiang"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

### Due date

Due: October 5 at 11:59pm. 

### Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0       | 20        |
| Problem 1       | --        |
| Problem 2       | 40        |
| Problem 3       | 40        |
| Optional survey | No points |


### Problem 0

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. We first load the libraries needed for this assignment.

```{r load_libraries}
library(tidyverse)
library(readxl)
```


### Problem 1

Since the solution for Problem 1 has been provided, the work for this problem will not be showcased in this submitted document.



### Problem 2

This code chunk imports and  from `Trash Wheel Collection Data.xlsx`. Using arguments in read_excel, we specify the "Mr. Trash Wheel" sheet and omit non-data entries (i.e., rows with notes/figures and columns containing notes), namely, the first row with the figures and texts; and the 15th and 16th columns.

```{r}
mr_trash_wheel = read_excel("./data/Trash Wheel Collection Data.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N550")
```

The next code chunk cleans up variable names such that they are reasonable in context, omits rows that do not include dumpster-specific data, rounds the number of sports balls to the nearest integer and converts the result to an integer variable, and convert the variable "year" from lists to doubles/numbers.

```{r}
mr_trash_wheel = 
  janitor::clean_names(mr_trash_wheel) %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(sports_balls),
         year = as.numeric(year)) 
```

Similarly, we import and tidy the data set for the sheet "Professor Trash Wheels": we import the data, omit rows and columns with notes and figures, cleans variable names, and omit the non-dumpster-specific rows.

```{r}
prof_trash_wheel = 
  read_excel("./data/Trash Wheel Collection Data.xlsx", 
             sheet = "Professor Trash Wheel", 
             range = "A2:M97") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster)
```

Now, we combine the two data sets into a single tidy data set. We first add an additional variable to both data sets to specify Trash Wheel.

```{r}
mr_trash_wheel$trash_wheel <- "Mr"
prof_trash_wheel$trash_wheel <- "Professor"
```

At last, we combine the two data sets.

```{r}
trash_wheel_data <- full_join(mr_trash_wheel, prof_trash_wheel) 
```

Finally, we provide a description for the joined data set. The number of observations in the resulting data set is `r nrow(trash_wheel_data)`. There are `r ncol(trash_wheel_data)` variables in this data set, and for some of the key variables are year (e.g., `r trash_wheel_data$year[1]`),... From the available data, the total weight of trash collected by Professor Trash Wheel was `r sum(trash_wheel_data$weight_tons[which(trash_wheel_data$trash_wheel=='Professor')])` tons, and the total number of sports balls collected by Mr. Trash Wheel in 2020 was r`sum(trash_wheel_data$sports_balls[which(trash_wheel_data$year==2020)], na.rm = TRUE)`. 



### Problem 3

Firstly, we import and clean the data in pols-month.csv. We then break the date variable "mon" into integer variables year, month, and day; replace month number with month name, create a variable "president" that specifies whether the the president is GOP or Dem, and remove the "day" variable.

```{r}
pols_month <- read_csv("./data/pols-month.csv") %>% 
  separate(col = mon, into = c("year", "month", "day")) %>% 
  mutate(year = as.numeric(year), 
         month = month.name[as.numeric(month)], 
         day = as.numeric(day), 
         president = ifelse(prez_gop == 1, "gop", "dem")) %>%
  select(-day, -prez_gop, -prez_dem) %>%
  janitor::clean_names()
```

We now turn to import and clean the data set of snp.csv. We break the "date" variable into integer variables year, month, and day; replace month number with month name, remove the day variable, arrange according to year and month, and organize such that year and month are leading columns.

```{r}
snp <- read_csv("./data/snp.csv") %>% 
  separate(col = date, into = c("month", "day", "year")) %>%
  mutate(year = as.numeric(year) + 2000, # to keep consistency with pols_month
         month = month.name[as.numeric(month)],
         day = as.numeric(day)) %>%
  select(-day) %>%
  relocate(year) %>% 
  janitor::clean_names()
```

Next, we import and tidy the unemployment data such that it has only three variables: year, month, and unemployment rate.

```{r}
unemployment <- read_csv("./data/unemployment.csv") %>%
  pivot_longer(Jan:Dec, 
               names_to = "month", 
               values_to = "unemployment") %>%
  mutate(month = month.name[match(month, month.abb)]) %>%
  janitor::clean_names()
```

Finally, we merge the three data sets by first merging snp into pols, and then merging unemployment into the result.

```{r}
pols_data <- left_join(left_join(pols_month, snp), unemployment)
```

Now we provide a description for the separate data sets and the final resulted joined data set. The pols_month data set contains `r nrow(pols_month)` observations and `r ncol(pols_month)` variables, including year, month, the number of GOP/Dem governors, senators, and representatives, respectively, and if the president of the particular month is Democrat or Republican. On the other hand, the snp data set contains `r nrow(snp)` observations and only `r ncol(snp)` variables; namely, year, month, and the closing values of the S&P stock index on some day during the given year and month. Finally, the unemployment data set contains `r nrow(unemployment)` and `r ncol(unemployment)` variables: year, month, and the percentage of unemployment date of a given year and month. At last, the result data set from joining the three data sets, pols_data, contains `r ncol(pols_data)` variables and `r nrow(pols_data)` observations. Some of the key variable names are `r colnames(pols_data)[1]`, `r colnames(pols_data)[2]`, `r colnames(pols_data)[9]`, `r colnames(pols_data)[10]`, and `r colnames(pols_data)[11]`. Furthermore, the range is between `r min(pols_data$year)` and `r max(pols_data$year)` with a difference of `r max(pols_data$year) - min(pols_data$year)` years.